SDSS-IV MaNGA: The Radial Profile of Enhanced Star Formation in Close Galaxy Pairs

### 2020-12-16 maintex.tex
text revisions

\usepackage{epstopdf}

@@

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 175 star-forming galaxies in close pairs with those of mass-matched isolated galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts.
\end{abstract}

@@

Among the 6142 MaNGA IFUs considered here, 404 have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in \S~\ref{sec:data}. These requirements reduce our sample down to 87 star forming MaNGA targets with inside-IFU companions. While the companions to the inside-IFU MaNGA targets contained are within the MaNGA IFUs, we only use the MaNGA targets when building radial profiles. We leave out the identified companions from the analysis because the MaNGA targets are covered by the NSA catalog with geometries and total stellar masses while the companions are not covered by the NSA catalog.

@@

To show how the population of galaxy pairs differs from isolated galaxies, we will compare our pair sample to a sample of isolated galaxies in the MaNGA survey. We build this control sample from the MaNGA target galaxies which have no spectroscopic companions within $r_{\rm p}$ $<$ 50 kpc and $\Delta v$ $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a control sample of 1888 star forming control galaxies.

It has been shown that the SFR in galaxies is dependent on both the stellar mass and the redshift of the galaxies \citep[e.g.,][]{Noeske:2007}. We will compare our interacting galaxies with isolated galaxies of a similar stellar mass and redshift, to account for these other dependencies. To do this, we will use two different methods of pair - control comparison.

@@

We fit the geometry with apertures from both catalogs when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA apertures are compared against controls fitted with the NSA apertures and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures. Further, when defining a galaxy's geometry with the \citet{Simard:2011} apertures we also use the masses given in the catalog for the total stellar mass of the galaxy. Note that although we carefully treat the two catalogs separately whenever possible, we find excellent agreement in the geometry parameters of isolated galaxies in both catalogs.

@@

Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. We build an azimuthally averaged radial profile for each MaNGA star forming galaxy and later stack and differentiate these profiles in the subsequent analysis.

@@

\begin{deluxetable*}{lcccccccccc}

\tablestypesize{\small} \tablewidth{0pt}
\tablecaption{Pair Catalog}
\centering
\tablehead{\colhead{Plate} & \colhead{IFU} & \colhead{Sample} & \colhead{\Delta\theta} & \colhead{r_{\rm p}} & \colhead{\Delta \rm{log}(M)} & \colhead{R.A.} & \colhead{Decl.} & \colhead{Redshift} & \colhead{log(M)} \\
\colhead{} & \colhead{} & \colhead{} & \colhead{(arcseconds)} & \colhead{(kpc)} & \colhead{(dex)} & \colhead{(deg)} & \colhead{(deg)} & \colhead{} & \colhead{log(M_\odot)} \\
\colhead{(1)} & \colhead{(2)} & \colhead{(3)} & \colhead{(4)} & \colhead{(5)} & \colhead{(6)} & \colhead{(7)} & \colhead{(8)} & \colhead{(9)} & \colhead{(10)}}

\startdata
%\begin{tabular}{cccccccccc}
%Plate & IFU & Sample & Sep_arc & Sep_kpc & d_logm & RAdeg & Decdeg & z & logm \\
7443 & 12703 & 1 & 7.3 & 5.8 & 0.0 & 229.52558 & 42.74584 & 0.0 & 10.8 \\
7960 & 12703 & 1 & 78.2 & 39.0 & 1.2 & 259.22831 & 31.52755 & 0.0 & 10.1 \\
7964 & 3704 & 1 & 55.0 & 19.7 & 1.7 & 316.25170 & -0.93328 & 0.0 & 9.2 \\
7975 & 12704 & 2 & 2.9 & 4.9 & 0.6 & 324.58642 & 11.34867 & 0.1 & 11.1 \\
7977 & 9102 & 2 & 8.6 & 10.5 & 2.1 & 332.83066 & 12.18472 & 0.1 & 10.8 \\
8078 & 6104 & 2 & 30.1 & 26.2 & -0.1 & 42.73943 & 0.36941 & 0.0 & 10.1 \\
8078 & 12703 & 1 & 35.4 & 20.0 & 0.9 & 42.57027 & 0.09199 & 0.0 & 10.8 \\
8079 & 1902 & 1 & 7.3 & 3.1 & 1.5 & 42.54777 & -0.99910 & 0.0 & 9.5 \\
8079 & 12702 & 1 & 61.7 & 27.4 & -0.6 & 41.86858 & -0.26820 & 0.0 & 10.3 \\
8080 & 9101 & 2 & 11.0 & 9.7 & 0.0 & 47.34003 & -0.40794 & 0.0 & 9.9 \\
%\end{tabular}
%\end{table*}

\enddata
\end{deluxetable*}

>>>>

%\usepackage{epstopdf}

@@

\begin{abstract}
We compare the radial profiles of the specific star formation rate (sSFR) in a sample of 175 star-forming galaxies in close pairs with those of mass-matched \textbf{control} galaxies in the SDSS-IV MaNGA survey. We find that the sSFR is centrally enhanced (within one effective radius) in interacting galaxies by $\sim$0.25 dex and that there is a weak sSFR suppression in the outskirts of the galaxies of $\sim$0.1 dex. We stack the differences profiles for galaxies in five stellar mass bins between \logm\ $=$ 9.0$-$11.5 and find that the sSFR enhancement has no dependence on the stellar mass. The same result is obtained when the comparison galaxies are matched to each paired galaxy in both stellar mass and redshift. In addition, we find that that the sSFR enhancement is elevated in pairs with nearly equal masses and closer projected separations, in agreement with previous work based on single-fiber spectroscopy. We also find that the sSFR offsets in the outskirts of the paired galaxies are dependent on whether the galaxy is the more massive or less massive companion in the pair. The more massive companion experiences zero to a positive sSFR enhancement while the less massive companion experiences sSFR suppression in their outskirts.
\end{abstract}

@@

\textbf{Broad-line AGN comprise $\sim$0.9\% of MaNGA's galaxy sample where \citet{Lacerda:2020} found a fraction of $\sim$2.8\% from the CALIFA survey and \citet{Sanchez:2018} found a fraction of $\sim$1.33\% from the MaNGA's MPL-5 sample (which observed $\sim$2700 galaxies).  
}

Among the 6142 MaNGA IFUs considered here, 404 have multiple galaxy objects within the IFU. We further restrict the sample by setting a relative velocity cut of $\Delta v$ $<$ 500 km s$^{-1}$ to remove projected companions. Given the redshift range of the MaNGA sample and the size of MaNGA's IFUs, the maximum projected separation for a companion galaxy in the IFU is $\sim$40 kpc. Again, the galaxy sample is also restricted to galaxies with a stellar mass range of \logm\ $=$ 9.0$-$11.5 and galaxies in the Primary or Secondary MaNGA subsamples. We also require that the galaxies are classified as star forming as described in \S~\ref{sec:data}. These requirements reduce our sample down to 87 star forming MaNGA targets with inside-IFU companions. While the companions to the inside-IFU MaNGA targets contained are within the MaNGA IFUs, we only use the MaNGA targets when building radial profiles. We leave out the identified companions from the analysis because the MaNGA targets are covered by the NSA catalog with geometries and total stellar masses while the companions are not covered by the NSA catalog. \textbf{While we use the identified companions to select galaxy pairs, we only build radial profiles for the MaNGA target galaxies. We do this because the identified companions are not included in the NSA catalog which we use to build the radial profiles for the galaxies.}

@@

To show how the population of galaxy pairs differs from isolated galaxies, we will compare our pair sample to a sample of \textbf{control} galaxies in the MaNGA survey. We build this control sample from the MaNGA target galaxies which have no spectroscopic companions within $r_{\rm p}$ $<$ 50 kpc and $\Delta v$ $<$ 500 km s$^{-1}$ either inside or outside of the IFU. This gives us a control sample of 1888 star forming control galaxies.

It has been shown that the SFR in galaxies is dependent on both the stellar mass and the redshift of the galaxies \citep[e.g.,][]{Noeske:2007}. We will compare our interacting galaxies with \textbf{control} galaxies of a similar stellar mass and redshift, to account for these other dependencies. To do this, we will use two different methods of pair - control comparison.

@@

We fit the geometry with apertures from both catalogs when available. When using the first method of pair-control comparison, where pairs and controls are grouped into stellar mass bins, we fit the control galaxies with the NSA apertures. When making the comparison with the second method, paired galaxies fitted with the NSA apertures are compared against controls fitted with the NSA apertures and paired galaxies fitted with the \citet{Simard:2011} apertures are compared against controls fitted with the \citet{Simard:2011} apertures. Further, when defining a galaxy's geometry with the \citet{Simard:2011} apertures we also use the masses given in the catalog for the total stellar mass of the galaxy. Note that although we carefully treat the two catalogs separately whenever possible, we find excellent agreement in the geometry parameters of \textbf{control} galaxies in both catalogs.

@@

Within each radius bin we take the median of the specific star formation rate. The profiles' errors are the standard error of the mean of the data within each radius bin. \textbf{Here we define the standard error of the mean, $\sigma_{\overline{x}}$, as,
\begin{equation}
\sigma_{\overline{x}} = \frac{\sigma}{\sqrt{n}},
\end{equation}
where $\sigma$ is the standard deviation and $n$ is the sample size.} We build an azimuthally averaged radial profile for each MaNGA star forming galaxy and later stack and differentiate these profiles in the subsequent analysis.

@@

\begin{deluxetable*}{cccccccccc}
%\tablestypesize{\small} \tablewidth{0pt}
\tablecaption{Pair Catalog}
\tablehead{
\colhead{Plate} & \colhead{IFU} & \colhead{Sample} & \colhead{\Delta\theta} & \colhead{r_{\rm p}} & \colhead{\Delta \rm{log}(M)} & \colhead{R.A.} & \colhead{Decl.} & \colhead{Redshift} & \colhead{log(M)} \\
\colhead{} & \colhead{} & \colhead{} & \colhead{(arcseconds)} & \colhead{(kpc)} & \colhead{(dex)} & \colhead{(deg)} & \colhead{(deg)} & \colhead{} & \colhead{log(M_\odot)} \\
\colhead{(1)} & \colhead{(2)} & \colhead{(3)} & \colhead{(4)} & \colhead{(5)} & \colhead{(6)} & \colhead{(7)} & \colhead{(8)} & \colhead{(9)} & \colhead{(10)}
}
\decimals
\startdata
7443 & 12703 & 1 & 7.3 & 5.8 & 0.0 & 229.52558 & 42.74584 & 0.04027 & 10.8 \\
7960 & 12703 & 1 & 78.2 & 39.0 & 1.2 & 259.22831 & 31.52755 & 0.02475 & 10.1 \\
7964 & 3704 & 1 & 55.0 & 19.7 & 1.7 & 316.25170 & -0.93328 & 0.01759 & 9.2 \\
7975 & 12704 & 2 & 2.9 & 4.9 & 0.6 & 324.58642 & 11.34867 & 0.08886 & 11.1 \\
7977 & 9102 & 2 & 8.6 & 10.5 & 2.1 & 332.83066 & 12.18472 & 0.06309 & 10.8 \\
8078 & 6104 & 2 & 30.1 & 26.2 & -0.1 & 42.73943 & 0.36941 & 0.04421 & 10.1 \\
8078 & 12703 & 1 & 35.4 & 20.0 & 0.9 & 42.57027 & 0.09199 & 0.02808 & 10.8 \\
8079 & 1902 & 1 & 7.3 & 3.1 & 1.5 & 42.54777 & -0.99910 & 0.02129 & 9.5 \\
8079 & 12702 & 1 & 61.7 & 27.4 & -0.6 & 41.86858 & -0.26820 & 0.02198 & 10.3 \\
8080 & 9101 & 2 & 11.0 & 9.7 & 0.0 & 47.34003 & -0.40794 & 0.04482 & 9.9 \\
\enddata
\end{deluxetable*}
